The ChEMBL database as Linked Open Data

### 2012-03-06 article.tex
Added Valery and the RSC affiliation details, emailed by Antony
Full names of the people in the Kasabi team
Added new lines (making easier-to-read git diffs) and s/Chembl/ChEBML/

% For one author:
\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
% Two or more authors (order to be determined):
\author[B]{\fnms{Ola} \snm{Spjuth}},
\author[C]{\fnms{Peter} \snm{Ansell}},
\author[D]{\fnms{Antony} \snm{Williams}},
\author[E]{\fnms{Janna} \snm{Hastings}},
\author[F]{\fnms{John} \snm{Overington}},
\author[F]{\fnms{Anna} \snm{Gaulton}},

\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands\\
E-mail: egon.willighagen@maastrichtuniversity.nl}
\runningauthor{E.L. Willighagen et al.}
\address[B]{Uppsala University}
\address[C]{University of Queensland}
\address[D]{xxx}
\address[E]{ChEBI, European Bioinformatics Institute}
\address[F]{ChEMBL, European Bioinformatics Institute}

@@

The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases \cite{Belleau2008}. A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}. These links are vital to provide context for use cases that require a correlation between chemical structures and other scientific data.

The Linked Data server used by Bio2RDF has been reconfigured for Chembl to provide URL based services for standard URI resolution, along with text and link searches \footnote{\url{https://github.com/ansell/chembl-webapp}}. It proxies the standard item URIs by translating URLs between those requested by users and the URIs that are available in the databases. For example, if the Chembl web application is running on the users local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}} will be resolved from the database using the full original URI \footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}. If the user requested an RDF document, using content negotiation, the original URIs will be unchanged, however, if the user requested an HTML document, the results will contain both the original RDF triples, represented using RDFa, with links that resolve using the users local machine.

The links services enables the Chembl application to derive both forward links, originating in Chembl \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}}, and backward links, originating in other databases, such as LODD, Bio2RDF \footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF. These services are vital to efficiently navigate the Linked Data web, as it is both impractical and inefficient to require users to crawl the web before they can discover all of the relevant resources. These services are currently only supplied as web services from Chembl and Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data providers in the future. Datasets that are available in SPARQL endpoints can be queried for links efficiently using templated queries in the Chembl web application.

@@

Tim and Foo of Kasabi. What S.H. Name of the Bioinformatics department at Uppsala University.

>>>>

% For one author:
\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
% Two or more authors (order to be determined):
\author[B]{\fnms{Ola} \snm{Spjuth}},
\author[C]{\fnms{Peter} \snm{Ansell}},
\author[D]{\fnms{Antony} \snm{Williams}},
\author[D]{\fnms{Valery} \snm{Tkachenko}},
\author[E]{\fnms{Janna} \snm{Hastings}},
\author[F]{\fnms{John} \snm{Overington}},
\author[F]{\fnms{Anna} \snm{Gaulton}},

\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands\\
E-mail: egon.willighagen@maastrichtuniversity.nl}
\runningauthor{E.L. Willighagen et al.}
\address[B]{Uppsala University}
\address[C]{University of Queensland}
\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
\address[E]{ChEBI, European Bioinformatics Institute}
\address[F]{ChEMBL, European Bioinformatics Institute}

@@

The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data
server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
A number of these databases are referenced in ChEMBL, including ChEBI, Pubmed, and both the
Uniprot protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
vital to provide context for use cases that require a correlation between chemical
structures and other scientific data.

The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
based services for standard URI resolution, along with text and link searches
\footnote{\url{https://github.com/ansell/chembl-webapp}}. It proxies the standard item
URIs by translating URLs between those requested by users and the URIs that are available
in the databases. For example, if the ChEMBL web application is running on the users
local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the
article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}}
will be resolved from the database using the full original URI
\footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}. If the
user requested an RDF document, using content negotiation, the original URIs will be unchanged,
however, if the user requested an HTML document, the results will contain both the
original RDF triples, represented using RDFa, with links that resolve using the users local machine.

The links services enables the Chembl application to derive both forward links, originating in
ChEMBL \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}},
and backward links, originating in other databases, such as LODD, Bio2RDF
\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
These services are vital to efficiently navigate the Linked Data web, as it is both
impractical and inefficient to require users to crawl the web before they can discover all of the
relevant resources. These services are currently only supplied as web services from ChEMBL and
Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data
providers in the future. Datasets that are available in SPARQL endpoints can be queried for
links efficiently using templated queries in the ChEMBL web application.

@@

Tim Hodson and Zach Beauvais of Kasabi for their support. What S.H. Name of the Bioinformatics department at Uppsala University.
